---
title: "Hypothesis testing of connectome distirbution difference"
author: "Zhaoqi Liu"
date: "5/6/2021"
output: html_document
---

```{r,message=FALSE}
library(R.matlab)
library(plot3D)
library(rgl) #produce interactive 3-D plots.
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r}
#load data
load("~/Downloads/STOR893/FinalProject/STOR-893-Final-Project/Data/merge_df.RData")
```


We want to assess whether there are significant difference in distribution of brain PC scores among subjects having major depression symptom versus subjects not having major depression. Since there are only 95 subjects have diagnosed major depression, we choose all subjects with major depression symptom and randomly choose 95 healthy subjects as two groups. 
We used Maximum Mean Discrepancy (MMD) test to obtain p values for
differences in the brain connectomes across the two groups for trait 130.

```{r,message=FALSE}
if(!require(kernlab)) { install.packages("kernlab", repos = "http://cran.us.r-project.org"); library("kernlab") }
```

```{r}
df_dep<-df[,c(1,2:121,251:252)]
df_dep<-na.omit(df_dep) #1028 obs
df_dep_sorted<-df_dep[order(df_dep$Trait_130),]
df_dep_sorted$depression<-as.factor(ifelse(df_dep_sorted$Trait_130==1,0,1))
```

```{r}
set.seed(893)
df_dep_yes<-tail(df_dep_sorted,nrow(df_dep_sorted[df_dep_sorted$depression==1,])) # group with major depression 
random.sample<-sample(1:nrow(df_dep_sorted[df_dep_sorted$depression==0,]),size=nrow(df_dep_sorted[df_dep_sorted$depression==1,]))
df_dep_no<-df_dep_sorted[random.sample,] # healthy group 
```


## two sample t test for PCA1 
```{r}
t.test(df_dep_yes$FCfeature_1,df_dep_no$FCfeature_1,var.equal = FALSE)

#SC
t.test(df_dep_yes$SCfeature_1,df_dep_no$SCfeature_1,var.equal = FALSE)
```

## MMD
### For FC 

```{r}
fc.yes<-as.matrix(df_dep_yes[,2:61])
fc.no<-as.matrix(df_dep_no[,2:61])
kmmd(fc.yes,fc.no,kernel="rbfdot",alpha=0.05,asymptotic=TRUE)
```



### For SC
```{r}
sc.yes<-as.matrix(df_dep_yes[,62:121])
sc.no<-as.matrix(df_dep_no[,62:121])
kmmd(sc.yes,sc.no,kernel="rbfdot")
```

The test results show that for both SC and FC, we don't have evidence to say that the distribution of brain PC score are different between healthy group and the group with diagnosed major depression. We cannot use Hotelling test because the PCA scores are not Gaussian multivariate-distributed. 


### Data Visualization 

```{r, webgl=TRUE}
df_dep_subset<-rbind(df_dep_yes,df_dep_no)

FC.PC1<-df_dep_subset$FCfeature_1
FC.PC2<-df_dep_subset$FCfeature_2
FC.PC3<-df_dep_subset$FCfeature_3

summary(df_dep_subset[,2:4])
#interactive 3d plot 
#green means healthy, red means with major depression 
plot3d(FC.PC1,FC.PC2,FC.PC3,type="p",size=3,col=ifelse(df_dep_subset$depression==0,"#1B9E77","#D95F02"))
```

```{r}
scatter3D(FC.PC1,FC.PC2,FC.PC3,colvar=as.numeric(df_dep_subset$depression),
          bty="b2",col=c("#1B9E77","#D95F02"),colkey=FALSE,
          pch=19,cex=0.5,theta=-60,phi=20,
          main= "FC", xlab= "PC1", ylab="PC2",zlab="PC3")
text3D(c(0.05,0.06),c(-0.08,-0.075),c(0.02,0.04),labels=c("Healthy","Major Depression"),add=TRUE,cex=0.8)
points3D(c(0.05,0.06),c(-0.075,-0.07),c(0.02,0.04),col=c("#1B9E77","#D95F02"),colkey=FALSE,add=TRUE,pch=19,cex=0.5)
```


```{r,webgl=hook_webgl}
SC.PC1<-df_dep_subset$SCfeature_1
SC.PC2<-df_dep_subset$SCfeature_2
SC.PC3<-df_dep_subset$SCfeature_3
summary(df_dep_subset[,62:64])

##green means healthy, red means with major depression 
plot3d(SC.PC1,SC.PC2,SC.PC3,type="p",size=3,col=ifelse(df_dep_subset$depression==0,"#1B9E77","#D95F02"))

```


```{r}
scatter3D(SC.PC1,SC.PC2,SC.PC3,colvar=as.numeric(df_dep_subset$depression),
          bty="b2",col=c("#1B9E77","#D95F02"),colkey=FALSE,
          pch=19,cex=0.5,theta=-60,phi=20,
          main= "SC", xlab= "PC1", ylab="PC2",zlab="PC3")
text3D(c(0.07,0.08),c(-0.11,-0.105),c(0.005,0.025),labels=c("Healthy","Major Depression"),add=TRUE,cex=0.8)
points3D(c(0.07,0.08),c(-0.105,-0.1),c(0.005,0.025),col=c("#1B9E77","#D95F02"),colkey=FALSE,add=TRUE,pch=19,cex=0.5)
```


